# Parsing HTML, XML, and JSON files using R
## Data 607 assignment
### Heather Geiger; March 18, 2018

### Introduction

Instructions:

"Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more
than one author. For each book, include the title, authors, and two or three other attributes that you find
interesting.

Take the information that you’ve selected about these three books, and separately create three files which
store the book’s information in HTML (using an html table), XML, and JSON formats.

Write R code, using your packages of choice, to load the information from each of the three sources into
separate R data frames. Are the three data frames identical?"

For this assignment, I chose three different cookbooks from Food Network authors.

HTML, XML, and JSON files are all available here:

https://github.com/heathergeiger/Parsing_HTML_XML_and_JSON

### Parsing and comparing results

Load libraries.

```{r, echo=TRUE,eval=FALSE}
library(XML)
library(RJSONIO)
```

```{r, echo=FALSE,eval=TRUE}
#Load Rdata file.
#This contains the three data frames, plus json_parsed_original as output by fromJSON command.
#The steps below were all actually run, but on a different server using R instead of Rstudio. So, I load in the results from Rdata file, and then show code but mostly do not actually evaluate it below.
load("parsed_data_frames.Rdata")
```

Parse XML file.

```{r, echo=TRUE,eval=FALSE}
xml_parsed <- xmlParse("books.xml")
root <- xmlRoot(xml_parsed)
books_from_xml_dat <- xmlToDataFrame(root)
```

Parse HTML file.

```{r, echo=TRUE,eval=FALSE}
html_parsed <- readHTMLTable("books_table.html")
books_from_html_dat <- html_parsed[[1]]
```

Parse JSON file.

Start by looking at number of elements in JSON file, then that determines which elements to get to obtain the info data frame.

Save this in object json_parsed.

```{r, echo=TRUE,eval=FALSE}
json_parsed_original <- fromJSON("books_table.json")
length(json_parsed_original)
length(names(json_parsed_original[[1]]))
json_parsed <- json_parsed_original[[1]][[1]]
```

```{r, echo=FALSE,eval=TRUE}
length(json_parsed_original)
length(names(json_parsed_original[[1]]))
```

Take the number of items in the first element of json_parsed as the number of columns, and the number of elements in json_parsed as the number of rows.

Then, make an empty data frame.

```{r, echo=TRUE,eval=FALSE}
books_from_json_dat <- data.frame(matrix(NA,
				nrow=length(json_parsed),
				ncol=length(json_parsed[[1]]),
				dimnames=list(1:length(json_parsed),
					names(json_parsed[[1]]))),
			stringsAsFactors=FALSE)
```

Fill the empty data frame row by row with the data from the JSON.

```{r, echo=TRUE,eval=FALSE}
for(i in 1:length(json_parsed))
{
books_from_json_dat[i,] <- unlist(json_parsed[[i]])
}
```

Check what data frames look like, and if they are actually the same.

```{r}
books_from_html_dat
books_from_xml_dat
books_from_json_dat

identical(books_from_html_dat,books_from_xml_dat)
identical(books_from_html_dat,books_from_json_dat)

class(books_from_html_dat$Author)
class(books_from_json_dat$Author)
```

We find that HTML and XML produce the exact same result. JSON results look the same, but columns are character instead of factor as in HTML/XML.
